---
title: 'Using aRules on the Titanic dataset '
author: 'Aymane Elbernoussi'
date: '4/08/22'
output: html_document
---

---

# Instructions
Using the Titanic dataset to explore descriptive statistics, functions, and
association rules. Download the titanic dataset `titanic.raw.rdata` from one
of two locations:

- https://github.com/ethen8181/machine-learning/tree/master/association_rule/R
- https://sites.google.com/a/rdatamining.com/www/data/titanic.raw.rdata?attredirects=1
 

---

```{r setup, message = FALSE}
# Add your library below.
#install.packages("arules")
library(arules)
#install.packages("arulesViz")
library(arulesViz)
library(dplyr)

```

# Step 0 - Load the data
Using the instructions above, load the dataset and save it as `t`.

# Step 1 - Descriptive stats (0.5 point for each answer)

1. Compute the percentage of people that survived.
2. Compute the percentage of people that were children.
3. Compute the percentage of people that were female.
4. Finally, compute the percentage of people that were in first class.

```{r, "Step 1"}
load("titanic.raw.rdata")
t <- titanic.raw

ttotal <- nrow(t)
survived <- filter(t,Survived == "Yes")

percent_survived <- nrow(survived)/ttotal

child <- nrow(filter(t,Age == "Child"))
percent_isChild <- child/ttotal

female <- nrow(filter(t,Sex == "Female"))
percent_isFemale <- female/ttotal

firstclass <- nrow(filter(t,Class == "1st"))
percent_infirstclass <- firstclass/ttotal


```

---

# Step 2 - More descriptive stats (0.5 point for each answer)


1. What percentage of children survived? Your answer should be written such as # 13.75% of children survived
2. What percentage of female survived?
3. What percentage of first-class people survived?
4. What percentage of third-class people survived?

```{r, "Step 2"}
# Write your code below.
child_survived <- nrow(filter(survived,Age == "Child"))
percent_childSurvived <- child_survived/nrow(filter(t, Age == "Child"))
#Out of all children, 52% survived

female_survived <- nrow(filter(survived,Sex == "Female"))
percent_femaleSurvived <- female_survived/nrow(filter(t, Sex == "Female"))
#Out of all females, 73% survived 

firstclass_survived <- nrow(filter(survived,Class == "1st"))
percent_firstclassSurvived <- firstclass_survived/nrow(filter(t, Class == "1st"))
#Out of all first class members, 62% survived 

thirdclass_survived <- nrow(filter(survived,Class == "3rd"))
percent_thirdclassSurvived <- thirdclass_survived/nrow(filter(t, Class == "3rd"))
#Out of all third class members, 25% survived

```

---

# Step 3 - Writing a function (0.5 point for each answer)

## Step 3.1 - Function 1
Write a function that returns a new dataframe of people that satisfies the 
specified criteria of sex, age, class and survived as parameters. I'm giving you 
the answer for this question:

```
```

```{r, "Step 3.1"}
# Write your code below.
myfunction1 <- function(a,b,c,d){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
df4 <- df3[df3$Survived == d,] # filter the data that satisfied the criteria that "Survived" = d
return(df4)}
# test the function with a sample data
myfunction1("1st","Female","Adult","No")


```

## Step 3.2 - Function 2
Write a function, using the previous function, that calculates the percentage 
(who lives, who dies) for a specified (parameters) of class, gender and age 
considering the entire number of data. The function passes four arguments. 
Include the following code properly in your function by improvising names of objects.

```
p <- nrow(df)/nrow(t) # calculate the percentage
```

```{r, "Step 3.2"}
# Write your code below.
myfunction2 <- function(a,b,c,d){
df1 <- t[t$Class == a,] # filter the data that satisfied the criteria that "Class" = a
df2 <- df1[df1$Sex == b,] # filter the data that satisfied the criteria that "Sex" = b
df3 <- df2[df2$Age == c,] # filter the data that satisfied the criteria that "Age" = c
df4 <- df3[df3$Survived == d,] # filter the data that satisfied the criteria that "Survived" = d
survived <- (nrow(df4)/nrow(t))*100 
return(survived)}
```

## Step 3.3 - Use the function (male)
Use the function to compare age and third-class male survival rates. 

```{r, "Step 3.3"}
# Write your code below.
myfunction2("3rd","Male","Adult","Yes")
# 3.4%
myfunction2("3rd","Male","Child","Yes")
# 0.59% 
```
People in which category are more likely to survive? 

> [Adult males in third class were more likely to survive compared to male children in 3rd class]


## Step 3.4 - Use the function (female)
Use the function to compare age and first-class female survival rates.

```{r, "Step 3.4"}
# Write your code below.
myfunction2("1st","Female","Adult","Yes")
# 6.36%
myfunction2("1st","Female","Child","Yes")
# 0.04% 

```
People in which category are more likely to survive? 

> [Adult Females in first class were more likely to survive compared to female children in 1st class]

---

# Step 4 - Use aRules (0.5 point for each answer)

1. Use aRules to calculate some rules (clusters) for the titanic dataset.
2. Visualize the results.

```{r, "Step 4.1"}
# Write your code below.
ruleset <- apriori(t,parameter = list(support = 0.01, confidence = 0.35))
plot(ruleset)
inspect(ruleset)
goodrules <- ruleset[quality(ruleset)$lift>3.3]
inspect(goodrules)
plot(goodrules)
```



